HFI in RISC-V

### 2025-05-01 paper.tex
Updated Paper

\section{Background}
\subsection{Software-based Fault Isolation}
Modern applications consistently rely on a large number of various libraries for key functionality, however, this leads to a problem as bugs within these libraries can become a vector of attack for the entire application.
To mitigate this, it is critical to contain the scope of what each library can access or affect.
Traditionally, general isolation was considered at the granularity of processes, containers, and virtual machines, which do achieve the functionality of isolation of libraries through hardware-level protection via the MMU and user-kernel privileges.
Although such these isolation methods are functional, they have significant performance overhead from the context switching and memory management. 
Due to the fact that the executing libraries are quite small, it motivates isolation that is fine-grained and lightweight.

@@

Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions. Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
In order to implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_enter} has funct3 $= 001$. Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.

@@

\subsection{Implementation - Explicit Regions}
Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExpllicitDataRegion} to store metadata about explicit regions. These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support. While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction. This difference is due to the ISA differences between x86 and RISC-V. Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.

For each of the load and store instructions, we would check the whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions). If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction. If the check fails, HFI would trap and prevent access to the address.

@@

\subsection{Evaluation - wasm2c}
We wanted to additionally add some of the evaluation done by the original HFI paper. We attempted to modify wasm2c by building on the work done already. The wasm2c code converts wasm files to C source and header and we wanted to modify it to use our modified h-prefix load and store instructions. We were unable to complete this step unfortunately due to some issues with building and running the module. However, this doesn't detract from the implementation of HFI itself. Although it would have been valuable to have used our HFI implementation in wasm2c, our unit tests still comprehensively demonstrate that HFI is feasible and functional in RISC-V. Our implementation is able to achieve the sandboxing functionality we wanted and effectively shows that HFI can extend to other architectures beyond x86. Hence, our work sufficiently shows the feasibility of HFI in RISC-V.

@@

\subsection{Expand Number of Explicit Regions}
Previously mentioned, we currently implemented only a singular explict region with regards to the number of instructions that we would introduce.
In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in Section~\ref{sec:approach}, we only have a finite number of available opcodes.

@@

\subsection{Cycle Accurate Simulator}
A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in a cycle accurate simulator, such as gem5, as done in the original HFI paper \cite{HFI}.
Since we used QEMU, an environment where the serialization of \textit{hfi\_enter} and \textit{hfi\_exit} would not produce observable effects, we chose not to implement serialization for Spectre mitigation.
In contrast, a cycle-accurate simulator enables observation of microarchitectural behavior, allowing for both functional correctness validation and evaluation of speculative execution resistance.
Moreover, with accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
As an implementation in a cycle accurate simulator is comparatively much slower in code modification and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.

>>>>

\section{Background}
\subsection{Software-based Fault Isolation}
Modern applications consistently rely on a large number of various libraries for key functionality, however, this leads to a problem as bugs within these libraries can become a vector of attack for the entire application.
To mitigate this, it is critical to contain the scope of what each library can access or affect.
Traditionally, general isolation was considered at the granularity of processes, containers, and virtual machines, which do achieve the functionality of isolation of libraries through hardware-level protection via the MMU and user-kernel privileges.
Although these isolation methods are functional, they have significant performance overhead from the context switching and memory management. 
Due to the fact that the executing libraries are quite small, it motivates isolation that is fine-grained and lightweight.

@@

Our work then began by creating a basic internal register called \textit{hfi\_status}, which serves as a flag bit for verifying whether HFI can be toggled on or off via our newly implemented instructions. Through the use of QEMU logging we display the value of \textit{hfi\_status} and use it as a validator that the instructions have run.
For our first instruction, we have \textit{hfi\_enter}, which sets \textit{hfi\_status} $=1$ when executed. 
Additionally, we have \textit{hfi\_exit}, which sets \textit{hfi\_status} $=0$.
RISC-V contains unmapped opcodes that can be used for the additional of new instructions.
In order to implement our instruction we have opted to use the free opcode $0001011$.
We are then able to differentiate between the two functions by modifying the funct3 value of the instructions such that \textit{hfi\_enter} has funct3 $= 000$ and \textit{hfi\_exit} has funct3 $= 001$. Adding the control and status register (CSR) \textit{hfi\_status} and implementing the \textit{hfi\_enter} and \textit{hfi\_exit} gave us basic functionality to enable and disable HFI.

@@

\subsection{Implementation - Explicit Regions}
Next, we moved on to implement explicit regions by defining a new struct \textit{HFIExplicitDataRegion} to store metadata about explicit regions. These structs were also stored in a new array in the CPU Architecture State of length based on the global variable indicating the number of regions we wanted to support. While the original HFI paper had the h-prefix \textit{mov} instruct (\textit{hmov}), we had h-prefix load and store instruction. This difference is due to the ISA differences between x86 and RISC-V. Hence, for our implementation, we focused on providing support for a single region and implemented the following instructions: \textit{hlb0}, \textit{hlh0}, \textit{hlw0}, \textit{hlbu0}, \textit{hlhu0}, \textit{hlwu0}, \textit{hld0}, \textit{hsb0}, \textit{hsh0}, \textit{hsw0}, and \textit{hsd0}.

For each of the load and store instructions, we would check whether both the beginning of the access address and end of the address access fall within a given data region (region 0 for the h-prefix and 0-postfix instructions). If HFI is enabled and the load or store fall within a properly configured region, then the load or store will be allowed through and proceed to load or store the given address by passing on the access addresses to the corresponding non-HFI load or store instruction. If the check fails, HFI would trap and prevent access to the address.

@@

\subsection{Evaluation - wasm2c}
We wanted to additionally add some of the evaluation done by the original HFI paper. We attempted to modify wasm2c by building on the work done already. The wasm2c code converts wasm files to C source and header and we wanted to modify it to use our modified h-prefix load and store instructions. Unfortunately, we were unable to complete this due to some issues with building and running the module. However, this doesn't detract from the implementation of HFI itself. Although it would have been valuable to have used our HFI implementation in wasm2c, our unit tests still comprehensively demonstrate that HFI is feasible and functional in RISC-V. Our implementation is able to achieve the sandboxing functionality we wanted and effectively shows that HFI can extend to other architectures beyond x86. Hence, our work sufficiently shows the feasibility of HFI in RISC-V.

@@

\subsection{Expand Number of Explicit Regions}
Previously mentioned, we currently implemented only a singular explicit region with regards to the number of instructions that we would introduce.
In the simplest of implementations, we would just associate a separate instruction for each explicit region, but as we have discussed in Section~\ref{sec:approach}, we only have a finite number of available opcodes.

@@

\subsection{Cycle Accurate Simulator}
A more comprehensive implementation for evaluation of security and performance in addition to ISA semantics can be considered in a cycle accurate simulator, such as gem5, as done in the original HFI paper \cite{HFI}.
Since we used QEMU, an environment where the serialization of \textit{hfi\_enter} and \textit{hfi\_exit} would not produce observable effects, we chose not to implement serialization for Spectre mitigation.
In contrast, a cycle-accurate simulator enables observation of microarchitectural behavior, allowing for both functional correctness validation and evaluation of speculative execution resistance.
Moreover, with an accurate cycle simulator, the performance overhead for HFI implementation can be validated when utilized within a SFI system.
As an implementation in a cycle accurate simulator is comparatively much slower in code modification and execution as compared to QEMU, as a feasibility metric for correctness, an implementation in QEMU was ideal with respect to our circumstances.
